---
title: "analysis"
author: "Jia Qi Beh"
date: "2025-08-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Global OptrA manuscript - analysis

This document contains all R packages and codes used in the bioinformatic analyses of the manuscript 'Global dissemination of optrA-mediated linezolid resistance
in enterococci'.


# Libraries

```{r}
library(tidyverse)
library(ggforce)
```


# Genome download

Download Illumina reads from NCBI SRA using fastq-dump
```{bash}
fastq-dump --outdir <OutputFolder> --gzip --skip-technical <AccessionNumber>
```

Download assemblies from NCBI using ncbi-entrez
```{bash}
# Create accessions.txt file
for line in $(cat accessions.txt); do efetch -db nucleotide -id ${line} -format fasta > ${line}.fasta ; done
```


# fastp
version 0.23.2

```{bash}
# create accessions.txt
# run fastp
for line in $(cat accessions.txt); do fastp --thread 16 -r -W 4 -M 20 -l 36 -i ${line}_1.trim.fastq.gz -I ${line}_2.trim.fastq.gz -o ${line}_R1_trimmed.fastq.gz -O ${line}_R2_trimmed.fastq.gz -j ${line}.json -h ${line}.html ; done
```


# short-read assembly
shovill version 1.1.0

```{bash}
# create accession.txt
# run shovill
while read line; do nice shovill --outdir ${line} --R1 ${line}_R1_trimmed.fastq.gz --R2 ${line}_R2_trimmed.fastq.gz --cpus 16; done < accessions.txt
```


# quast
version 5.2.0

```{bash}
# create accession.txt
# run quast
for line in $(cat accessions.txt); do ${line}/contigs.fa -o ${line} -t 16; done
```


# abritamr
version 1.0.11

```{bash}
abritamr run --contigs input.txt --species Enterococcus_faecium -j 16
abritamr run --contigs input.txt --species Enterococcus_faecalis -j 16
```


# fastani
version 1.33

```{bash}
for line in $(cat accessions.txt); do /home/fastani/FastANI-master/fastANI -q ${line}.fasta -r reference.fasta -o ${line}.out ; done
```


# mlst
version 2.23.0

```{bash}
mlst --scheme efaecium assembly.fa --threads 10
```


# snippy + iqtree
version 4.4.5

```{bash}
# run snippy
snippy --cpus 16 --outdir output --ref reference.gb --R1 ID_R1.trimmed.fastq.gz --R2 ID_R2.trimmed.fastq.gz

# run snippy-core
snippy-core AUSMDU* --ref reference.gbff

# run iqtree v2.0.6
iqtree -s core.aln -m GTR+I+G -B 1000 -nt 16
```


# cd-hit-est
version 4.8.1

```{bash}
cd-hit -i input.fasta -o output -c 0.95 -b 20 -n 5 -M 1100 -T 5
```


# mob-typer
mobsuite version 3.0.3

```{bash}
mob_typer --infile assembly.fa --out_file output.txt
```


# mashtree
version 1.4.6

```{bash}
mashtree --mindepth 0 --numcpus 2 --outmatrix mash_matrix.txt *.fasta > mashtree.dnd
```


# contig-puller
version 2.0

```{bash}
for line in $(cat accessions.txt); do contig-puller.py --db optrA_reference.fasta --out ${line}.fasta --id 80 --cpus 16 ${line}.fasta; done
```


# panaroo

```{bash}
# copy all .gff files into input_gff/
# run panaroo with default core threshold 0.95
conda activate panaroo
panaroo -i input_gff/*.gff -o results_threshold95 --clean-mode strict -t 30 --remove-invalid-genes

# Gene graph neighbourhood
panaroo-gene-neighbourhood --gene optrA --graph final_graph.gml --out neighbourhood.txt --expand_no 2
```


# minimap2 + samtools
minimap2 version 2.17
samtools version 1.21

```{bash}
# make assembly accessions.txt
# run minimap2
for line in $(cat accessions.list); do /home/minimap2/minimap2 -a plasmid_seq.fasta ${line}.fasta > ${line}.sam ; done

# run samtools
for line in $(cat accessions.list); do samtools view -S -b ${line}.sam > ${line}.bam; done
for line in $(cat accessions.list); do samtools sort -o ${line}.sorted.bam ${line}.bam ; done
for line in $(cat accessions.list); do samtools index ${line}.sorted.bam; done

# Generate depth information
for line in $(cat accessions.list); do samtools depth -a ${line}.sorted.bam > ${line}.depth.txt; done

# count total nt covered in the reference Tn6674
total_positions=57536  # reference length
while IFS= read -r accession; do
    samtools depth -a ${accession}.sorted.bam > ${accession}.depth.txt
    covered_positions=$(awk '$3 > 0 {count++} END {print count}' ${accession}.depth.txt)
    coverage_percentage=$(echo "scale=2; $covered_positions / $total_positions * 100" | bc)
    echo "Coverage percentage for ${accession}: $coverage_percentage%"
done < accessions.list > output.txt
```


# clinker

```{bash}
clinker *.gbff -ufo -j 12 -o output -p plot.html
```

