---
title: "visualisation"
author: "Jia Qi Beh"
date: "2025-08-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Global OptrA manuscript - visualisation

This document contains all R packages and codes used to create the figures in the manuscript 'Global dissemination of optrA-mediated linezolid resistance in enterococci'.


# Libraries

```{r}
pacman::p_load(
  tidyverse,     # data management and visualization
  rio,
  UpSetR,        # special package for combination plots
  ggupset,
  ComplexUpset,
  readxl,
  readr,
  reshape2,      # melt() function
  svglite,
  hrbrthemes,
  viridis,
  RColorBrewer,
  pheatmap,
  cowplot,
  patchwork,
  micropan,
  matrixStats,
  ggforce, 
  reshape2,
  ggtree,
  here,
  ape,
  treeio,
  ggnewscale,
  phytools,
  phangorn,
  pals,
  scales,
  lubridate,
  rworldmap,
  cowplot, 
  stats,
  gplot)       # special package for combination plots
```

Set working directory

```{r}
setwd("~/Downloads/")
```


# Figure 1

Manually constructed.


# Figure 2

## 2a - map

Import data
```{r}
data <- read.csv("figure_2/map_metadata.csv")
data
```

Count the total number of isolates for each country
```{r}
country_counts <- data %>% dplyr::count(country)
total_isolates <- sum(country_counts$n)

# Convert counts to percentages
country_counts <- country_counts %>%
  mutate(percentage_category = cut(
    n, 
    breaks = c(-Inf, 20, 40, 60, 80, Inf), 
    labels = c("<20%", "20-40%", "40-60%", "60-80%", ">80%"), 
    include.lowest = TRUE
  ))
country_counts
```

Plot map
```{r}
# Define color palette
colours <- c('#caf0f8', '#90e0ef','#00b4d8', '#0077b6', '#03045e') #own colour palette

# Make plot
pdf("figure_2a.pdf", width=80, height=60)
sPDF <- joinCountryData2Map(country_counts, joinCode='NAME', nameJoinColumn='country')
map.allOptrA <- mapCountryData(
  sPDF, 
  nameColumnToPlot='percentage_category', 
  catMethod="categorical", 
  missingCountryCol='dark grey', 
  colourPalette=brewer.pal(5, "Blues"),  # Use a color palette for categories
  addLegend=TRUE
)
dev.off()
```


## 2b - optrA Efs

Count each source type for E. faecalis
```{r}
data <- read.csv("figure_2/optrAEfs.csv")
data %>% dplyr::count(sample_type) %>%
  arrange(desc(n))
```

Make plot
```{r}
plot <- 
  ggplot(data, aes(x = year, fill= sample_type)) + 
  geom_bar(color = "black", width = 0.5) +
  xlab("Collection year")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", 
                                    margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("animal"="#A8D0DB", "clinical/human"="#E49273", 
                               "environment"="#A37A74", na.value = "white"))+
  theme_bw()
plot

ggsave("figure_2b.pdf", plot, width = 8, height = 6)
```

## 2c - optrA Efm

Count each source type for E. faecium
```{r}
data <- read.csv("figure_2/optrAEfm.csv")
data %>% dplyr::count(sample_type) %>%
  arrange(desc(n))
```

Make plot
```{r}
plot <- 
  ggplot(data, aes(x = year, fill= sample_type)) + 
  geom_bar(color = "black", width = 0.5) +
  xlab("Collection year")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", 
                                    margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("animal"="#A8D0DB", "clinical/human"="#E49273", 
                               "environment"="#A37A74", na.value = "white"))+
  theme_bw()
plot

ggsave("figure_2c.pdf", plot, width = 8, height = 6)
```


# Figure 3

## 3a - mash tree

Clean up abritamr output
```{r}
# Read file
df <- read.csv("figure_3/abritamr.csv")
df

# Create a new column to show optrA presence/absence
sample_data <- df %>%
  mutate(optrA = if_else(
    if_any(everything(), ~ grepl("optrA", ., fixed = TRUE)),
    "yes",
    "no"
  ))
sample_data

# Count number of isolates positive for optrA (optrA == yes)
sum(sample_data$optrA == "yes") # 55 isolates
```

Make tree
```{r}
# Load tree file
iqtree <- ape::read.tree("figure_3/mashtree.dnd")
iqtree

# Plot tree with rhierBAPS levels
tree.1 <- ggtree(iqtree)
tree.1

# Check whether the Isolate IDs in both files align correctly
sample_data$Isolate %in% iqtree$tip.label
iqtree$tip.label %in% sample_data$Isolate

setdiff(sample_data$Isolate, iqtree$tip.label) 
setdiff(iqtree$tip.label, sample_data$Isolate) 

# Add root and scale bar
tree_mid <- midpoint(iqtree)
tree_mid
ape::write.tree(tree_mid, file='~/Downloads/midpoint_tree.nwk')
```

Make plot
```{r}
# Plot tree
tree.1 <- ggtree(iqtree)
tree.1
tree.2 <- tree.1 %<+% sample_data
tree.3 <- tree.2 + geom_tippoint(size=0.5, aes(color = optrA))

# Set manual tip colours
tree.3 <- tree.2 + geom_tippoint(size=0.5, aes(color = optrA)) +
  scale_color_manual(values = c("yes" = "#fb8500", "no" = "#7F7F7F")) +
  theme(legend.position = "right")
tree.3

ggsave("figure_3a.pdf", tree.3, width=8, height=6)
```

## 3b - dendrogram

Import data and make dendrogram
```{r}
# Import data
df <- read.csv("figure_3/transposed_metadata.csv")

# Distance matrix
d <- dist(df)

# Hierarchical clustering
hc <- hclust(d, method = "ward.D2")
print(hc)

# Make dendrogram
pdf("figure_3b.pdf")
plot(as.dendrogram(hc), main = "Efm dendrogram", horiz = TRUE)
  # horiz = rotate plot horizontally (by default the dendrogram is vertical)
dev.off()
```

Draw heatmap for gene presence-absence matrix
```{r}
heatmap_df <- read.csv("figure_3/heatmap_metadata.csv")
binary_matrix <- as.matrix(heatmap_df[, ])

pdf("heatmap.pdf")
heatmap(binary_matrix, Rowv = NA, Colv = NA, col = c("white", "black"), 
        scale = "none", xlab = "Features", ylab = "Samples")
dev.off()
```

Other heatmaps were added manually.


# Figure 4

## 4a - mashtree

Clean up abritamr output
```{r}
# Read file
df <- read.csv("figure_4/abritamr.csv")
df

# Create a new column to show optrA presence/absence
sample_data <- df %>%
  mutate(optrA = if_else(
    if_any(everything(), ~ grepl("optrA", ., fixed = TRUE)),
    "yes",
    "no"
  ))
sample_data

# Count number of isolates positive for optrA (optrA == yes)
sum(sample_data$optrA == "yes") # 31 isolates
```

Make tree
```{r}
# Load tree file
iqtree <- ape::read.tree("figure_4/mashtree.dnd")
iqtree

# Plot tree with rhierBAPS levels
tree.1 <- ggtree(iqtree)
tree.1

# Check whether the Isolate IDs in both files align correctly
sample_data$Isolate %in% iqtree$tip.label
iqtree$tip.label %in% sample_data$Isolate

setdiff(sample_data$Isolate, iqtree$tip.label) 
setdiff(iqtree$tip.label, sample_data$Isolate) 

# Add root and scale bar
tree_mid <- midpoint(iqtree)
tree_mid
ape::write.tree(tree_mid, file='~/Downloads/midpoint_tree.nwk')
```

Make plot
```{r}
# Plot tree
tree.1 <- ggtree(iqtree)
tree.1
tree.2 <- tree.1 %<+% sample_data
tree.3 <- tree.2 + geom_tippoint(size=0.5, aes(color = optrA))

# Set manual tip colours
tree.3 <- tree.2 + geom_tippoint(size=0.5, aes(color = optrA)) +
  scale_color_manual(values = c("yes" = "#fb8500", "no" = "#7F7F7F")) +
  theme(legend.position = "right")
tree.3

ggsave("figure_4a.pdf", tree.3, width=8, height=6)
```

## 4b - dendrogram

Import data and make dendrogram
```{r}
# Import data
df <- read.csv("figure_4/transposed_metadata.csv")

# Distance matrix
d <- dist(df)

# Hierarchical clustering
hc <- hclust(d, method = "ward.D2")
print(hc)

# Make dendrogram
pdf("~/Downloads/dendrogram.pdf")
plot(as.dendrogram(hc), main = "Efm dendrogram", horiz = TRUE)
  # horiz = rotate plot horizontally (by default the dendrogram is vertical)
dev.off()
```

Draw heatmap for gene presence-absence matrix
```{r}
heatmap_df <- read.csv("figure_4/heatmap_metadata.csv")
binary_matrix <- as.matrix(heatmap_df[, ])

pdf("heatmap.pdf")
heatmap(binary_matrix, Rowv = NA, Colv = NA, col = c("white", "black"), 
        scale = "none", xlab = "Features", ylab = "Samples")
dev.off()
```

Other heatmaps were added manually.


# Figure 5

Pie charts for each country were created separately and added to map
```{r}
data <- data.frame(category = c("A", "B", "C"), 
                   value = c(0, 1, 0))

custom_colors <- c("A" = "#2a6f97", "B" = "#e29578", "C" = "#edf6f9")

p <- ggplot(data, aes(x = "", y = value, fill = category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = custom_colors) +
  theme_void() +
  theme(legend.position = "none")
p

ggsave("country_name.pdf", p, width = 6, height = 4, dpi = 300)
```


# Figure 6

Figure 6 was created using Clinker.


# Figure S1

Import tree
```{r}
tree <- ape::read.tree("figure_s1/core.aln.treefile")
tree
```

Add metadata
```{r}
sample_data <- read.csv("figure_s1/efs_metadata.csv")
View(sample_data)

# Check whether the Isolate IDs in both files align correctly
sample_data$ID %in% tree$tip.label
tree$tip.label %in% sample_data$ID 

setdiff(sample_data$ID, tree$tip.label) #metadata ID that are not in tree 
setdiff(tree$tip.label, sample_data$ID) #tree ID that are not in metadata 

# Add midpoint root
tree_mid <- midpoint(tree)
tree_mid
```

Plot tree
```{r}
tree.1 <- ggtree(tree_mid)
tree.1

# Add root and scale bar
tree.2 <- 
  ggtree(tree_mid, branch.length = 0.5) %<+% sample_data +
  geom_rootedge(rootedge = 0.0005) +
  geom_treescale(x=-0.005, y=50)
tree.2
```

Heat map 1: Continent
```{r}
continent <- data.frame("continent" = sample_data[,c("continent")])
rownames(continent) <- sample_data$ID

tree.3 <- tree.2 + new_scale_fill()
h1 <- 
  gheatmap(tree.3, continent, width = 0.05, color="white", offset=0.01,
           colnames = FALSE) +
  scale_fill_manual(name = "continent",
                    values = c("#423E3B","#FF2E00","#FEA82F", 
                               "#FFFECB","#5448C8", "#2a9d8f", "#5448C8"),
                    na.value="white",  # set colour of NA values
                    breaks = c("Oceania", "North America", "Europe",
                               "Asia", "South America", "asia"),
                    labels = c("Oceania", "North America", "Europe", 
                               "Asia", "South America", "asia")) +
  theme(legend.position = "none")
h1
```

Heatmap 2: Source type
```{r}
sample_type <- data.frame("sample_type" = sample_data[,c("sample_type")])
rownames(sample_type) <- sample_data$ID

h2 <- h1 + new_scale_fill()
h3 <- 
  gheatmap(h2, sample_type, width = 0.05, color="white", offset=0.027,
           colnames = FALSE) +
  scale_fill_manual(name = "host",
                    values = c("#f28482", "#f28482", "#f28482", 
                               "#8ecae6", "#84a59d", "white"),
                    na.value="white",  # set colour of NA values
                    breaks = c("clinical/human", "human", "clinical", 
                               "animal", "environmental", "others"),
                    labels = c("clinical/human", "human", "clinical", 
                               "animal", "environmental", "others")) +
  theme(legend.position = "none")
h3
```

Heatmap 3: ST
Only top 10 STs annotated
```{r}
ST <- data.frame("ST" = sample_data[,c("ST")])
rownames(ST) <- sample_data$ID
sample_data$ST <- as.factor(sample_data$ST)  # convert into factor data type
sample_data <- sample_data %>%
  mutate(ST = case_when(
    ST %in% c('16','476','480','21','585','86','256', '632') ~ as.character(ST),
    TRUE ~ 'others'  # Convert everything else into 'others'
  ))

ST <- data.frame("MLST" = sample_data[,c("ST")])
rownames(ST) <- sample_data$ID

h4 <- h3 + new_scale_fill()
h5 <- 
  gheatmap(h4, ST, width = 0.05, color="white", offset=0.044,
           colnames = FALSE) +
  #colnames = TRUE, colnames_position= "top",
  #colnames_angle = 0, hjust = 0.5, 
  #font.size=3, colnames_offset_y = 1) +
  scale_fill_manual(name = "ST",
                    values = c("#588157", "#390099", "#BAF2D8",
                               "#F2BAC9","#669bbc","#F2E2BA",
                               "#C49799","#ee9b00", "white"), 
                    na.value="white",  # set colour of NA values
                    breaks = c('16','476','480','21','585','86','256', '632'),
                    labels = c('16','476','480','21','585','86','256', '632')) +
  theme(legend.position = 'bottom', 
        legend.key.size = unit(0.2, 'cm'))
h5
```

Heatmap 4: mlplasmids
```{r}
mlplasmids_prediction <- data.frame("mlplasmids_prediction" = sample_data[,c("mlplasmids_prediction")])
rownames(mlplasmids_prediction) <- sample_data$ID

h6 <- h5 + new_scale_fill()
h7 <- 
  gheatmap(h6, mlplasmids_prediction, width = 0.05, color="white", offset=0.061,
           colnames = FALSE) +
  scale_fill_manual(name = "mlplasmids_prediction",
                    values = c("#fca311","#a8dadc"),
                    na.value="white",  # set colour of NA values
                    breaks = c("Chromosome", "Plasmid"),
                    labels = c("Chromosome", "Plasmid")) +
  theme(legend.position = "none")
h7
```

Heatmap 5: optrA context
```{r}
sample_data$predicted_optrA_context <- as.factor(sample_data$predicted_optrA_context)

predicted_optrA_context <- data.frame("predicted_optrA_context" = sample_data[,c("predicted_optrA_context")])
rownames(predicted_optrA_context) <- sample_data$ID

h8 <- h7 + new_scale_fill()
h9 <- 
  gheatmap(h8, predicted_optrA_context, width = 0.05, color="white", offset=0.078,
           colnames = FALSE) +
  scale_fill_manual(name = "mlplasmids_prediction",
                    values = c("#9b5de5","#f15bb5", "#fee440"),
                    na.value="white",  # set colour of NA values
                    breaks = c("fexA-optrA", "fexA-optrA-ermA", "Tn6674"),
                    labels = c("fexA-optrA", "fexA-optrA-ermA", "Tn6674")) +
  theme(legend.position = "none")
h9
```

Heatmap 6: Tn6674 variant
```{r}
sample_data$Tn6674_variant <- as.factor(sample_data$Tn6674_variant)

Tn6674_variant <- data.frame("Tn6674_variant" = sample_data[,c("Tn6674_variant")])
rownames(Tn6674_variant) <- sample_data$ID

h10 <- h9 + new_scale_fill()
h11 <- 
  gheatmap(h10, Tn6674_variant, width = 0.05, color="white", offset=0.095,
           colnames = FALSE) +
  scale_fill_manual(name = "mlplasmids_prediction",
                    values = c("#a4c3b2","#e26d5c"),
                    na.value="white",  # set colour of NA values
                    breaks = c("WT", "truncated"),
                    labels = c("WT", "truncated")) +
  theme(legend.position = "none")
h11
```

Save plot
```{r}
ggsave("figure_s1.pdf", h11, width=8, height=6, dpi=300)
```


# Figure S2

Import tree
```{r}
tree <- ape::read.tree("figure_s2/core.aln.treefile")
tree
```

Add metadata
```{r}
sample_data <- read.csv("figure_s2/efm_metadata.csv")
View(sample_data)

# Check whether the Isolate IDs in both files align correctly
sample_data$ID %in% tree$tip.label
tree$tip.label %in% sample_data$ID 

setdiff(sample_data$ID, tree$tip.label) #metadata ID that are not in tree 
setdiff(tree$tip.label, sample_data$ID) #tree ID that are not in metadata 

# Add midpoint root
tree_mid <- midpoint(tree)
tree_mid
```

Plot tree
```{r}
tree.1 <- ggtree(tree_mid)
tree.1
```

Add root and scale bar
```{r}
tree.2 <- 
  ggtree(tree_mid, branch.length = 0.5) %<+% sample_data +
  geom_rootedge(rootedge = 0.0005) +   # add root 
  #geom_rootpoint() +
  geom_treescale(x=-0.005, y=50)
tree.2
```

Heatmap 1: Continent
```{r}
continent <- data.frame("continent" = sample_data[,c("continent")])
rownames(continent) <- sample_data$ID

tree.3 <- tree.2 + new_scale_fill()
h1 <- 
  gheatmap(tree.3, continent, width = 0.05, color="white", offset=0.02,
           colnames = FALSE) +
  scale_fill_manual(name = "continent",
                    values = c("#423E3B","#FF2E00","#FEA82F", 
                               "#FFFECB","#5448C8"),
                    na.value="white",  # set colour of NA values
                    breaks = c("Oceania", "North America", "Europe", "Asia"),
                    labels = c("Oceania", "North America", "Europe", "Asia")) +
  theme(legend.position = "none")
h1
```

Heatmap 2: Source
```{r}
sample_type <- data.frame("sample_type" = sample_data[,c("sample_type")])
rownames(sample_type) <- sample_data$ID

h2 <- h1 + new_scale_fill()
h3 <- 
  gheatmap(h2, sample_type, width = 0.05, color="white", offset=0.04,
           colnames = FALSE) +
  scale_fill_manual(name = "host",
                    values = c("#f28482", "#8ecae6", "#84a59d"),
                    na.value="white",  # set colour of NA values
                    breaks = c("clinical/human", "animal", "environment"),
                    labels = c("clinical/human", "animal", "environment")) +
  theme(legend.position = "none")
h3
```

Heat map 3: ST
Only top 10 ST were annotated
```{r}
ST <- data.frame("ST" = sample_data[,c("ST")])
rownames(ST) <- sample_data$ID
sample_data$ST <- as.factor(sample_data$ST)  # convert into factor data type
sample_data <- sample_data %>%
  mutate(ST = case_when(
    ST %in% c('80','22','132','324','184','32','650') ~ as.character(ST),
    TRUE ~ 'others'  # Convert everything else into 'others'
  ))

ST <- data.frame("MLST" = sample_data[,c("ST")])
rownames(ST) <- sample_data$ID

h4 <- h3 + new_scale_fill()
h5 <- 
  gheatmap(h4, ST, width = 0.05, color="white", offset=0.06,
           colnames = FALSE) +
  #colnames = TRUE, colnames_position= "top",
  #colnames_angle = 0, hjust = 0.5, 
  #font.size=3, colnames_offset_y = 1) +
  scale_fill_manual(name = "Sample source",
                    values = c("#588157", "#390099", "#BAF2D8",
                               "#F2BAC9","#669bbc",
                               "#F2E2BA","#C49799","#ee9b00"), 
                    na.value="white",  # set colour of NA values
                    breaks = c('80','22','132','324','184','32','650'),
                    labels = c('80','22','132','324','184','32','650')) +
  theme(legend.position = 'bottom', 
        legend.key.size = unit(0.2, 'cm'))
h5
```

Heatmap 4: CC17
```{r}
sample_data$CC <- as.factor(sample_data$CC)
CC <- data.frame("CC" = sample_data[,c("CC")])
rownames(CC) <- sample_data$ID

h6 <- h5 + new_scale_fill()
h7 <- 
  gheatmap(h6, CC, width = 0.05, color="white", offset=0.08,
           colnames = FALSE) +
  scale_fill_manual(name = "van_type",
                    values = c("#c1121f"),
                    na.value="white",  # set colour of NA values
                    breaks = c("17"),
                    labels = c("17")) +
  theme(legend.position = "none")
h7
```

Heatmap 5: van type
```{r}
van_operon <- data.frame("van_operon" = sample_data[,c("van_operon")])
rownames(van_operon) <- sample_data$ID

h8 <- h7 + new_scale_fill()
h9 <- 
  gheatmap(h8, van_operon, width = 0.05, color="white", offset=0.1,
           colnames = FALSE) +
  scale_fill_manual(name = "van_type",
                    values = c("#219ebc","#fb8500"),
                    na.value="white",  # set colour of NA values
                    breaks = c("vanA", "vanD"),
                    labels = c("vanA", "vanD")) +
  theme(legend.position = "none")
h9
```

Heatmap 6: mlplasmids
```{r}
mlplasmids_prediction <- data.frame("mlplasmids_prediction" = sample_data[,c("mlplasmids_prediction")])
rownames(mlplasmids_prediction) <- sample_data$ID

h10 <- h9 + new_scale_fill()
h11 <- 
  gheatmap(h10, mlplasmids_prediction, width = 0.05, color="white", offset=0.12,
           colnames = FALSE) +
  scale_fill_manual(name = "mlplasmids",
                    values = c("#fca311","#a8dadc"),
                    na.value="white",  # set colour of NA values
                    breaks = c("Plasmid", "Chromosome"),
                    labels = c("Plasmid", "Chromosome")) +
  theme(legend.position = "none")
h11
```

Heatmap 7: optrA context
```{r}
predicted_optrA_context <- data.frame("predicted_optrA_context" = sample_data[,c("predicted_optrA_context")])
rownames(predicted_optrA_context) <- sample_data$ID

h12 <- h11 + new_scale_fill()
h13 <- 
  gheatmap(h12, predicted_optrA_context, width = 0.05, color="white", offset=0.14,
           colnames = FALSE) +
  scale_fill_manual(name = "mlplasmids",
                    values = c("#9b5de5","#f15bb5", "#fee440"),
                    na.value="white",  # set colour of NA values
                    breaks = c("optrA-erm", "Tn6261", "fexA-optrA-ermA"),
                    labels = c("optrA-erm", "Tn6261", "fexA-optrA-ermA")) +
  theme(legend.position = "none")
h13
```

Save plot
```{r}
ggsave("figure_s2.pdf", h13, width=8, height=6, dpi=300)
```


# Figure S3

## S3a - Efs

Load data
```{r}
data <- read.csv("figure_s3/efs_data.csv")
data
```

Count optrA variants
```{r}
count_variant <- 
  data %>% dplyr::count(OptrA_variant) %>%
  arrange(desc(n))
count_variant
```

Then select only OptrA variants with (N >=5) and change others into "Others"
```{r}
data.1 <-
  data %>%
  mutate(
    new_OptrAvariant = case_when(
      OptrA_variant == "G40N, G393D" ~ "Others",
      OptrA_variant == "T112K, Y176D, T481P" ~ "Others",
      OptrA_variant == "D401E" ~ "Others",
      OptrA_variant == "T112K, S147L, Y176D, T481P" ~ "Others",
      OptrA_variant == "Y176D, G393D, I622M" ~ "Others",
      OptrA_variant == "truncated" ~ "Others",
      OptrA_variant == "G393D" ~ "Others",
      OptrA_variant == "G40N, I287K, I622M" ~ "Others",
      OptrA_variant == "G87D, Y176D, G393D" ~ "Others",
      OptrA_variant == "Q219H" ~ "Others",
      OptrA_variant == "S147F, I287K" ~ "Others",
      OptrA_variant == "Y176D, A350V, G393D" ~ "Others",
      OptrA_variant == "Y176D, G394S" ~ "Others",
      OptrA_variant == "MANY" ~ "Others",
      TRUE ~ OptrA_variant   # keep all other values unchanged
    )
  )
data.1
```

Make plot
```{r}
plot <- 
  ggplot(data.1, aes(x = year, fill= new_OptrAvariant)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("Collection year")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("wild type"="#606c38", 
                               "I104R, Y176D, E256K"="#283618", 
                               "Y176D, T481P"="#dda15e", 
                               "Y176D, G393D"="#2A9D8F", 
                               "T112K, Y176D"="#c1121f", 
                               "Y176D, I622M"="#780000",
                               "Y176D"="#E76F51", 
                               "Y176D, D247N, G393D, I622M"="#0077b6", 
                               "G40N, I287K, G393D"="#ffba08",
                               "Others"="white",
                               na.value = "white"))+
  theme_bw()
plot

ggsave("figure_s3a.pdf", plot, width = 8, height = 6)
```


## S3b - Efm

Load data
```{r}
data <- read.csv("figure_s3/efm_data.csv")
data
```

Count optrA variants
```{r}
count_variant <- 
  data %>% dplyr::count(OptrA_variant) %>%
  arrange(desc(n))
count_variant
```

Then select only OptrA variants with (N >=5) and change others into "Others"
```{r}
data.1 <-
  data %>% 
  mutate(new_OptrAvariant = case_when(
                           OptrA_variant == "Y176D, A176V, G393D" ~ "Others",
                           OptrA_variant == "Y176D, D247N, G393D, I622M" ~ "Others",
                           OptrA_variant == "Y176D, T481P" ~ "Others",
                           OptrA_variant == "T94K, Y176D" ~ "Others",
                           OptrA_variant == "I104R, Y176D, E256K" ~ "Others",
                           OptrA_variant == "S147F, Y176D, I287K, A350V, G393D, S411G, T481P" ~ "Others",
                           OptrA_variant == "Y176D, G393D, I622M" ~ "Others",
                           OptrA_variant == "Y176D, L213P, I622M" ~ "Others",
                           OptrA_variant == "MANY" ~ "Others",
                           OptrA_variant == "truncated" ~ "Others",
                           TRUE ~ OptrA_variant   # keep all other values unchanged
                           ))
data.1
```

Make plot
```{r}
plot <- 
  ggplot(data.1, aes(x = year, fill= OptrA_variant)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("Collection year")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("Y176D, G393D"="#2A9D8F", 
                               "Y176D, I622M"="#780000", 
                               "T113K, S147L, Y176D, T481P"="#a2d2ff", 
                               "wild type"="#606C38", 
                               "G40D, Y176D, G393D"="#E4C1F9", 
                               "Y176D"="#E76F51",
                               "T113K, S147L, Y176D, I287K"="#a7c957", 
                               "Others"="white",
                               na.value = "white"))+
  theme_bw()
plot

ggsave("figure_s3b.pdf", plot, width = 8, height = 6)
```


# Figure S4

Figure S4 was created using Clinker.


# Figure S5

Figure S5 was created using Clinker.


# Figure S6

Tn6674
```{r}
data <- read.csv("figure_s6/Tn6674_metadata.csv")
data

plot <- 
  ggplot(data, aes(x = ST, fill= Tn6674)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("OptrA variant")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("yes"="#c1121f", "no"="white"))+
  theme_bw()
plot

ggsave("plot_1.pdf", plot, width = 8, height = 6)
```

fexA-optrA-ermA
```{r}
data <- read.csv("figure_s6/plasmid_metadata.csv")
data

plot <- 
  ggplot(data, aes(x = ST, fill= plasmid)) + 
  geom_bar(color = "#023047", width = 0.5) +
  xlab("OptrA variant")+
  ylab("Number of isolates") +
  theme(axis.title.x = element_text(size = 15, colour = "black", margin = margin(10, 0, 10, 0), face = "bold"),
        axis.text.x = element_text(size = 15, color = "black",),
        axis.text.y = element_text(size = 15, color = "black"),
        axis.title.y = element_text(colour = "black", margin = margin(0, 10, 0, 10)))+
  scale_fill_manual(values = c("yes"="#0077B6", "no"="white"))+
  theme_bw()
plot

ggsave("plot_2.pdf", plot, width = 8, height = 6)
```

Then manually combine both plots.


# Figure S7

Figure S7 was created using Clinker.


# Figure S8

Figure S8 was created using Clinker.










# Figure S9

## S9a - efs optrA plasmid

Load data
```{r}
data <- read.csv("figure_s9/coverage_optrAplasmids.csv")
data

ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 0.5) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()

# Save plot into object
p <- ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 1) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()
p

ggsave("figure_s9a.pdf", p, width=8, height=4, dpi=300)
```

## S9b - efs optrA genomes

Load data
```{r}
data <- read.csv("figure_s9/coverage_optrAgenomes.csv")
data

ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 0.5) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()

# Save plot into object
p <- ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 1) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()
p

ggsave("figure_s9b.pdf", p, width=8, height=4, dpi=300)
```

Annotations were manually added.


# Figure S10

## S10a - efm optrA plasmids

Load data
```{r}
data <- read.csv("figure_s10/coverage_optrAplasmids.csv")
data

ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 0.5) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()

# Save plot into object
p <- ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 1) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()
p

ggsave("figure_s10a.pdf", p, width=8, height=4, dpi=300)
```

## S10b - efm optrA genomes

Load data
```{r}
data <- read.csv("figure_s10/coverage_optrAgenomes.csv")
data

ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 0.5) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()

# Save plot into object
p <- ggplot(data, aes(x = position, y = total_coverage)) +
  geom_line() +
  geom_area(fill = "black", alpha = 1) + # set area under graph color and transparency
  labs(x = "Position", y = "Total Coverage", title = "Coverage Distribution") +
  theme_minimal()
p

ggsave("figure_s10b.pdf", p, width=8, height=4, dpi=300)
```

Annotations were manually added.

